---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
library('asdetect')
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# asdetect

asdetect is used to detect abrupt shifts in time series, by searching for large changes in gradient over a small time 

## Installation

You can install asdetect using:

``` {r, eval=F}
#install.packages('devtools')
devtools::install_github("caboulton/asdetect")
```

## Usage

The below example shows how to the functions work on an example time series, using the as_detect() function:

```{r}
eg1 <- tip_eg() #create a time series with an abrupt shift in it
plot(eg1, type="l", xlab='Time', ylab='X')
detect1 <- as_detect(eg1$x, dt=0.5) #run the detection algorithm
plot(detect1, type="l", xlab='Time', ylab='Detection Value', ylim=c(-1,1))
```

A time series of white noise will have no detection of an abrupt shift:

```{r}
eg2 <- white_noise_eg() #create a time series of white noise
plot(eg2, type="l", xlab='Time', ylab='X')
detect2 <- as_detect(eg2$x, dt=0.5) #run the detection algorithm
plot(detect2, type="l", xlab='Time', ylab='Detection Value', ylim=c(-1,1))
```

Using both detection objects we can determine if a shift has occurred using where_as():

```{r}
print(where_as(detect1$detect, dt=0.5)) #finds the minima or maxima of the detection time series above the default 0.7 threshold
print(where_as(detect2$detect, dt=0.5)) #warning message displayed if none can be found
```

It is important to note that depending o whether or not 'dt' is specified, depends on how inputs are used in follow-on functions. Information is provided in the individual help files of each function.

asdetect also includes shift_type() which determines whether a detected abrupt shift is a true shift or whether it is a flat section of an otherwise increasing/decreasing time series. Again, full information is provided in the help file.

## References

Boulton C. A. & Lenton T. M. (2019) 'A method for detecting abrupt shifts in time series' (in prep.)
